### 2015-08-25 manuscript.tex
Add comments from Rasmus

\section*{Introduction}
There are many methods for inferring the presence of admixture, e.g. methods
using simple summary statistics detecting deviations from phylogenetic symmetry
\cite{reich2009reconstructing} [TODO(mason) Also cite patterson paper and paper
from Monty's group], methods estimating admixture proportions with programs such
as Structure \cite{pritchard2000inference}, Admixture \cite{alexander2009fast}
or RFmic [cite] [TODO comment: we might want to make sure to add more citations
here, for example all the ones from our previous papers and the ones from the
ALDER paper. People like to see themselves cited]. However, there has been less
research on estimating admixture times, possibly because such methods require
data which were unavailable until the advent of high-throughput next generation
sequencing. Some recently developed methods use the inferred local ancestry of
sequences to construct admixture tract length distributions [TODO cite Pool Nielsen,
Gravel and yourself]. Over time, recombination is expected to decrease the
average lengths of admixture tracts. The length distribution of admixture tracts
is therefore informative about the time since admixture.  Much of the theory
relating to tracts lengths is based on Fisher's famous theory of junctions
\cite{fisher1949theory} and subsequent work by Stam \cite{stam1980distribution}
and others [TODO add as many citations here as possible]. For example,
\cite{baird2003distribution} first discussed the length distribution of tracts
descended from a single ancestor. These results informed later analyses of
admixture tract length distribution, such as \cite{pool2009inference},
\cite{gravel2012population}, and \cite{liang2014lengths}.
\cite{gravel2012population} also implemented the software program TRACTS, which
estimates admixture histories by fitting the tract length distribution, obtained
by local ancestry inference, to a exponential approximation.

@@

[TODO Could we say anything here about Neanderthals in Asia?]

% You may title this section "Methods" or "Models".
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis"
\section{Model}
We use a random union of gametes admixture model as described in
\cite{liang2014understanding}, which is an extension of the mechanistic
admixture model formulated by \cite{verdu2011general}. In this model, two or
more source populations contribute migrants to form an admixed population
consisting of $2N$ haploid individuals. Each generation in the admixed
population is formed through the recombination of randomly selected individuals
from the previous generation, with some individuals potentially replaced by
migrants from the source populations. For simplicity, we consider a model with
only two source populations. Furthermore, the first source population only
contributes migrants in the founding generation, $T$. The second source
population contributes migrants in the founding generation and possibly in one
or more generations thereafter. In generation $i$, for $i=T-1,\dots,0$ (before
the present), a fraction $m_i$ of the admixed population is replaced by
individuals from the second source population.

\section{Linkage Disequilibrium and Local Ancestry}
ROLLOFF and ALDER use the standard two-locus measure of LD between a SNP at
positions $x$ and another SNP at position $y$, which is a genetic distance $d$
to the right,

\begin{align}
	D_2(d) = \text{cov}(H_x,H_y),
	\label{D2}
\end{align}

where $H_x$ and $H_y$ represent the haplotype or genotypes of an admixed
chromosome at positions $x$ and $y$. In the case of haplotype data, $H_{i,x}=1$
if the $i^\text{th}$ sample is carrying the derived allele at the SNP at
position $x$, and is otherwise 0. Alternatively, for genotype data, $H_{i,x}$
take on values from $\{0,1/2,1\}$ depending on the number of copies of the
derived allele the $i^\text{th}$ sample is carrying at SNP position $x$.
We consider an additional site at position $z$, which is located a further
genetic distance $d'$ to the right of $y$. The three-loci LD, as defined by as
defined by \cite{bennett1952theory} and \cite{slatkin1972treating}, is given by

@@

\section{Weighted Linkage Disequilibrium}
As \cite{loh2013inferring} noted, we cannot use the LD in the admixed population directly, because the allele frequency differences in the source populations can be of either sign. As in As \cite{loh2013inferring} , we solve this problem by computing the product of the values of the three-point linkage disequilibrium coefficient with the product of the allele frequency differences. Using equation \ref{local_ancestry} we obtain
$$
\delta_x\delta_y\delta_z D_3(d,d')=\delta_x^2\delta_y^2\delta_z^2\mathbb{E}[\text{cov}(A_x,A_y,A_z)],
$$
because the local ancestry in the admixed sample is independent of the allele frequencies in the admixed population. For inference purposes, we estimate this function by averaging over triples of SNPs which are separated by distances of approximately $d$ and $d'$. The LD term is estimated from the admixed population, while the $\delta$'s are estimated from reference populations which are closely related to the two source populations. We notice that both this approach, as well as the previous approaches (e.g., \cite{loh2013inferring} ), does not take genetic drift in the source populations after the time of admixture into account, i.e. there is an assumption of both this mehtod and previous methods that the allele frequencies in the ancestral source populations can be approximated well using the allele frequencies in the extant populations.

@@

When using only the admixed population itself as a reference population, the method described above will be biased if the same samples are used to estimate both the linkage disequilibrium coefficients and the weights ($\delta_x$, $\delta_y$, and $\delta_z$). We cannot efficiently compute a polyache statistics like \cite{loh2013inferring}. At the cost of some power, we instead adopt the approach of \cite{pickrell2012inference} and separate the admixed population into two equal-sized groups. We then use one group to estimate the weights, and the other group to estimate linkage disequilibrium coefficients, and vice versa. This gives gives two unbiased estimates for the numerator of $\hat{a}$, which we then average. [TODO Comment: I didn't understand how this works.  Coudl we five more detail?]
\subsection{Fitting the Two-Pulse Model}
We fit equation \ref{alder_2p} to the estimates of the weighted LD using non-linear least squares, with two modifications. We added a proportionality constant to account for the expected square allele frequency difference between the source populations. We also subtracted out an affine term in the weighted LD which is due to population substructure \cite{loh2013inferring}. We estimated this by computing the three-way covariance between triples of chromosomes. We use the jackknife to obtain confidence intervals for the resulting estimates by leaving out each chromosome in turn and refitting on the data for the remaining chromosomes.  [TODO Comment: again, we need more detail here. These are important parts of the manuscript]
\section{Simulations}
We used the program macs \cite{chen2009fast} to generate two source populations which diverged 4000 generations ago and a coalescent simulation to generate an admixed population from the two source populations according to two-pulse and constant admixture models. We sampled 50 diploid individuals from the admixed and two source populations, each consisting of 20 chromosomes of length 1 Morgan. The effective population size was $2N=1000$ for the admixed population and two source populations. Using a two pulse model, we varied the migration probabilities and timings for each pulse to examine the accuracy of equation \ref{alder_2p}. We also simulated data for a model with a constant rate of admixture each generation, and compared this to the predictions made by equation \ref{cov}.
\section{Data Set}
We computed the weighted LD for the Mexican and Columbia populations in the first phase of the 1000 Genomes data set. These consisted of 66 individuals from Los Angeles and 60 individuals from Medellin, respectively. We used the 88 Yoruba samples as the one reference population. We computed the weighted LD on the genotypes to avoid effects of phasing errors.
% Results and Discussion can be combined.
\section{Results and Discussion}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\subsection{Simulations}

@@

[TODO: the color coding should be labeled for all figures.  Also, the y and x-axes should be labeled]

@@

[TODO Comment: we need  figures showing the similar one-dimensional LD for comparison.  Maybe the first thing to discuss is to show how you can get approximately the same 1D curve for different models which produce different 2D surfaces]

@@

\bibliography{thesis}

>>>>

\section*{Introduction}
There are many methods for inferring the presence of admixture, e.g. methods using simple summary statistics detecting deviations from phylogenetic symmetry \cite{reich2009reconstructing} [Also cite patterson paper and paper from Monty's group] , methods estimating admixture proportions with programs such as Structure \cite{pritchard2000inference},  Admixture \cite{alexander2009fast} or RFmic [cite] [comment: we might want to make sure to add more citations here, for example all the ones from our previous papers and the ones from the ALDER paper.  People like to see themselves cited]. However, there has been less research on estimating admixture times, possibly because such methods require data which were unavailable until the advent of high-throughput next generation sequencing. Some recently developed methods use the inferred local ancestry of sequences to construct admixture tract length distributions [cite Pool&Nielsen, Gravel and yourself]. Over time, recombination is expected to decrease the average lengths of admixture tracts. The length distribution of admixture tracts is therefore informative about the time since admixture.  Much of the theory relating to tracts lengths is based on Fisher's famous theory of junctions \cite{fisher1949theory} and subsequent work by Stam \cite{stam1980distribution} and others [add as many citations here as possible]. For example, \cite{baird2003distribution} first discussed the length distribution of tracts descended from a single ancestor. These results informed later analyses of admixture tract length distribution, such as \cite{pool2009inference}, \cite{gravel2012population}, and \cite{liang2014lengths}. \cite{gravel2012population} also implemented the software program TRACTS, which estimates admixture histories by fitting the tract length distribution, obtained by local ancestry inference, to a exponential approximation.

@@

[Could we say anything here about Neanderthals in Asia?]

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section{Model}
We use a random union of gametes admixture model as described in \cite{liang2014understanding}, which is an extension of the mechanistic admixture model formulated by \cite{verdu2011general}. In this model, two (or more) source populations contribute migrants to form an admixed population consisting of $2N$ haploid individuals. Each generation in the admixed population is formed through the recombination of randomly selected individuals from the previous generation, with some individuals potentially replaced by migrants from the source populations. For simplicity, we consider a model with only two source populations. Furthermore, the first source population only contributes migrants in the founding generation, $T$. The second source population contributes migrants in the founding generation and possibly in one or more generations thereafter. In generation $i$, for $i=T-1,\dots,0$ (before the present), a fraction $m_i$ of the admixed population is replaced by individuals from the second source population.

\section{Linkage Disequilibrium and Local Ancestry}
ROLLOFF and ALDER use the standard two-locus measure of LD between a SNP at positions $x$ and another SNP at position $y$, which is a genetic distance $d$ to the right, 
\begin{align}
D_2(d) = \text{cov}(H_x,H_y),
\label{D2}
\end{align}
where $H_x$ and $H_y$ represent the haplotype or genotypes of an admixed chromosome at positions $x$ and $y$. In the case of haplotype data, $H_{i,x}=1$ if the $i^\text{th}$ sample is carrying the derived allele at the SNP at position $x$, and is otherwise 0. Alternatively, for genotype data, $H_{i,x}$ take on values from $\{0,1/2,1\}$ depending on the number of copies of the derived allele the $i^\text{th}$ sample is carrying at SNP position $x$. 
We consider an additional site at position $z$, which is located a further genetic distance $d'$ to the right of $y$. The three-loci LD, as defined by as defined by \cite{bennett1952theory} and \cite{slatkin1972treating}, is given by

@@

\section{Weighted Linkage Disequilibrium}
As \cite{loh2013inferring} noted, we cannot use the LD in the admixed population directly, because the allele frequency differences in the source populations can be of either sign. As in As \cite{loh2013inferring} , we solve this problem by computing the product of the values of the three-point linkage disequilibrium coefficient with the product of the allele frequency differences. Using equation \ref{local_ancestry} we obtain
$$
\left\delta_x\delta_y\delta_z D_3(d,d')\right=\delta_x^2\delta_y^2\delta_z^2\mathbb{E}[\text{cov}(A_x,A_y,A_z)],
$$
because the local ancestry in the admixed sample is independent of the allele frequencies in the admixed population. For inference purposes, we estimate this function by averaging over triples of SNPs which are separated by distances of approximately $d$ and $d'$. The LD term is estimated from the admixed population, while the $\delta$'s are estimated from reference populations which are closely related to the two source populations. We notice that both this approach, as well as the previous approaches (e.g., \cite{loh2013inferring} ), does not take genetic drift in the source populations after the time of admixture into account, i.e. there is an assumption of both this mehtod and previous methods that the allele frequencies in the ancestral source populations can be approximated well using the allele frequencies in the extant populations.

@@

When using only the admixed population itself as a reference population, the method described above will be biased if the same samples are used to estimate both the linkage disequilibrium coefficients and the weights ($\delta_x$, $\delta_y$, and $\delta_z$). We cannot efficiently compute a polyache statistics like \cite{loh2013inferring}. At the cost of some power, we instead adopt the approach of \cite{pickrell2012inference} and separate the admixed population into two equal-sized groups. We then use one group to estimate the weights, and the other group to estimate linkage disequilibrium coefficients, and vice versa. This gives gives two unbiased estimates for the numerator of $\hat{a}$, which we then average. [Comment: I didn't understand how this works.  Coudl we five more detail?]
\subsection{Fitting the Two-Pulse Model}
We fit equation \ref{alder_2p} to the estimates of the weighted LD using non-linear least squares, with two modifications. We added a proportionality constant to account for the expected square allele frequency difference between the source populations. We also subtracted out an affine term in the weighted LD which is due to population substructure \cite{loh2013inferring}. We estimated this by computing the three-way covariance between triples of chromosomes. We use the jackknife to obtain confidence intervals for the resulting estimates by leaving out each chromosome in turn and refitting on the data for the remaining chromosomes.  [Comment: again, we need more detail here. These are important parts of the manuscript]
\section{Simulations}
We used the program macs \cite{chen2009fast} to generate two source populations which diverged 4000 generations ago and a coalescent simulation to generate an admixed population from the two source populations according to two-pulse and constant admixture models. We sampled 50 diploid individuals from the admixed and two source populations, each consisting of 20 chromosomes of length 1 Morgan. The effective population size was $2N=1000$ for the admixed population and two source populations. Using a two pulse model, we varied the migration probabilities and timings for each pulse to examine the accuracy of equation \ref{alder_2p}. We also simulated data for a model with a constant rate of admixture each generation, and compared this to the predictions made by equation \ref{cov}.
\section{Data Set}
We computed the weighted LD for the Mexican and Columbia populations in the first phase of the 1000 Genomes data set. These consisted of 66 individuals from Los Angeles and 60 individuals from Medellin, respectively. We used the 88 Yoruba samples as the one reference population. We computed the weighted LD on the genotypes to avoid effects of phasing errors.
% Results and Discussion can be combined.
\section{Results and Discussion}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\subsection{Simulations}

@@

[Comment: the color coding should be labeled for all figures.  Also, the y and x-axes should be labeled]

@@

[Comment: we need  figures showing the similar one-dimensional LD for comparison.  Maybe the first thing to discuss is to show how you can get approximately the same 1D curve for different models which produce different 2D surfaces]

@@

\bibliographystyle{plos_template}
\bibliography{}
